NLIVis: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-22 paper/intro.tex
update paper structure

One recent advance in neural natural language process models is the introduction of attention mechanisms~\cite{VaswaniShazeerParmar2017} (see Section~\ref{sec:attention}). Intuitively, the \emph{attention} captures where in the input the model determine it should focus on. The attention often represents as weights that assigned to individual word or a pair of word (i.e., the alignment between words in different sentences).
%
There has been many speculation on how attention works in various models. In this work, by utilizing the the proposed system, we are able to perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow user to change prediction, where we apply an constrained optimization to find the least amount of parameter update to produce the user assign label. Another word, the user can explore how changes in sentence input, attention, or prediction affect each other. Such type of exploration is not only essential to interpret the language inference problem, but can also be easily generated to other NLP model utilized the attention mechanisms.
%
In addition, we also enhanced the standard visual encoding (e.g., as a bipartite graph or as an matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar guided simplification of the visual represent.
Finally, as discussed in Section~\ref{sec:caseStudy}, the ability to examine how attention corresponds to grammatical structure also help domain expert speculate the potential benefit of including sentence linguistic structure in the design of attention component of the model.

The key contributions of the proposed works is summarized as the following:
\begin{itemize}
    \item Help understand how the model fail by examining stability of the prediction and visual summarization of predictions;

@@

    \item Enable the interrogation of the relationship between input sentences, attention, and the prediction by interactively exploring how the change in one place affect other parts of the model.

\end{itemize}
%
% In addition,

>>>>

One recent advance in neural natural language process models is the introduction of attention mechanisms~\cite{VaswaniShazeerParmar2017} (see Section~\ref{sec:attention}). Intuitively, the \emph{attention} captures where in the input the model determine it should focus on. The attention often represents as weights that assigned to individual word or a pair of word (i.e., the alignment between words in different sentences).
%
There has been many speculation on how attention works in various models. In this work, by utilizing the the proposed system, we are able to perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow user to change prediction, where we apply an constrained optimization to find the least amount of parameter update to produce the user assign label. Another word, the user can explore how changes in sentence input, attention, or prediction affect each other. Such type of exploration is not only essential to interpret the language inference problem, but can also be easily generated to other NLP model utilized the attention mechanisms.
%
In addition, we also enhanced the standard visual encoding (e.g., as a bipartite graph or as an matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar guided simplification of the visual representation.
%
Finally, as discussed in Section~\ref{sec:caseStudy}, the ability to examine how attention corresponds to grammatical structure also help domain expert speculate the potential benefit of including sentence linguistic structure in the design of attention component of the model.

The key contributions of the proposed works is summarized as the following:
\begin{itemize}
    \item Identify the visualization tasks for making sense of the role of attention in natural langage inference (section~\ref{sec:task});

    \item Help understand how the model fail by examining stability of the prediction and visual summarization of predictions;

@@

    \item Enable the interrogation of the relationship between input sentences, attention, and the prediction by interactively exploring how the change in one place affect other parts of the model.
\end{itemize}
%
% In addition,

### 2018-03-22 paper/relatedWorks.tex
minor changs
update paper structure

\section{Related Works}

How to interpret classification task
\begin{itemize}
    \item What feature / dimension / part-of-the-input should change to
    alter the outcome
    \item What Examples can you provide that is similar to the current
    instance
\end{itemize}
\begin{itemize}
    \item Interacting with Predictions: Visual Inspection of Black-box Machine Learning Models,
    Use partial dependency (varying along one feature and observer how prediction changes)
    \item “Why Should I Trust You?” Explaining the Predictions of Any Classifier (LIME)
\end{itemize}

(Limitation of the state-of-the arts)
Model independent explanations is useful to gain intuition, but they lack the ability to link the external observation with inner states/mechanism of the model.

Depth learning vis
\begin{itemize}
    \item ActiVis: Visual Exploration of Industry-Scale Deep Neural Network Models
    \item Analyzing the Training Processes of Deep Generative Models
    \item Visual Diagnosis of Tree Boosting Methods
    \item TreePOD: Sensitivity-Aware Selection of Pareto-Optimal Decision Trees
    \item LSTMVis: A Tool for Visual Analysis of Hidden State Dynamics in Recurrent Neural Networks
    \item Do Convolutional Neural Networks learn Class Hierarchy?
    \item DeepEyes: Progressive Visual Analytics for Designing Deep Neural Networks
\end{itemize}

Less emphasis on where does the fail

>>>>

\section{Related Works}
With the increasing demands for model accountability (e.g., what is the evidence for making the decision) and model fairness (e.g., is the prediction affected by the bias in the training data),
the interpretability of the machine learning model held significant implications. As a result, a wealth of researches have be focused on the explanation and interpretation of deep neural network models from both machine learning as well as visualization community.

\noindent\textbf{Model Interpretability.}
What consistitue interpretation for a model learning model
% The the rising field of explainable AI

~\cite{Doshi-Velez2017}
Discussion the underspecification of what interprability means~\cite{Lipton2016}
How to interpret classification task

Many works have been dedicated to providing intuitive explanations for a given prediction, several~\cite{RibeiroSinghGuestrin2016, KrausePererNg2016} of which approach the problem from a model agnostic approach that made them applicable to different applications (i.e., by fitting a simplier linear model near the prediction of interests).
%
Despite being invaluable for the providing human understandable explanation, the lack of the ability to access and explore the internal states of the model, which is vitally important to make sense of why a model fail, limits their ability for more in depth exploration and evaluation of the model.
% (Limitation of the state-of-the arts)
% Model independent explanations is useful to gain intuition, but they lack the ability to link the external observation with inner states/mechanism of the model.

\noindent\textbf{Visual Interpration.}
Visual analytics has demonstrate its effectiveness in the exploration of complex relationship. The application~\cite{TzengMa2005} has long been adopted for a
~\cite{KrausePererNg2016}
Several previous works~\cite{BilalJourablooYe2018} have dedicate to analyzing errors made in classification

Depth learning vis
% \begin{itemize}
%     \item ActiVis: Visual Exploration of Industry-Scale Deep Neural Network Models
%     \item Analyzing the Training Processes of Deep Generative Models
%     \item Visual Diagnosis of Tree Boosting Methods
%     \item TreePOD: Sensitivity-Aware Selection of Pareto-Optimal Decision Trees
%     \item LSTMVis: A Tool for Visual Analysis of Hidden State Dynamics in Recurrent Neural Networks
%     \item Do Convolutional Neural Networks learn Class Hierarchy?
%     \item DeepEyes: Progressive Visual Analytics for Designing Deep Neural Networks
% \end{itemize}

\noindent\textbf{Visualization of Neural Linguistic Model.}
Compare to the multitude of researchs on making sense of convolutional neural networks, relatively less works has been dedicated to the neural linguistic model.
%
However, the previous works~\cite{KarpathyJohnson2015, LiChenHovy2015, StrobeltGehrmannPfister2018, LiuBremerJayaraman2018} not only demonstrated the benefit of visualization in understanding NLP models, but also illuminated many possibility with the application of visual analytics in understanding neural linguistic models.
% There is many unresolve challenge and the
The LSTMvis~\cite{StrobeltGehrmannPfister2018} work utilized a line plot style visual encoding to visualize the time varying hidden states of recurent neutural network. The word embedding visualization work~\cite{LiuBremerJayaraman2018} illustrate how semantic relationship can be recovered by linear projection in the high-dimensional word embedding space produced by word2vec~\cite{MikolovSutskeverChen2013} or glove~\cite{PenningtonSocherManning2014}.

### 2018-03-22 paper/task.tex
update paper structure

\section{Task Analysis}
Effective identify and itemize the visualization tasks and the corresponding goal is the first step in designing a useful visual representation.

>>>>

\section{Task Analysis}
\label{sec:task}
Effective identify and itemize the visualization tasks and the corresponding goal is the first step in designing a useful visual representation.
